Vragenlijst thesis evaluatie

### 2015-06-28 doc/template/chap-4.tex
delay methods aanvullen

Infection by worms start with dropping the work on the network. Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node. Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through phising or trojan horses. More important for determining our delay is the propagation strategy. Propagation is performed by determining the next nodes to spread the worm to. The following are common propagation methods: 
\begin{description}
%\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning.
%% random, localized, topological or hit-list scanning.
%\begin{itemize}
\item \textbf{Selective Random scanning:} Some worms propagate by using the method of selective random targeting IP addresses. The worm selects randomly a part of the selected address space instead of scanning the whole address space. The reserved address blocks and the unassigned addresses are excluded from the address space. The rate of success for randomly chosen IP addresses is very low, but it is easy to implement. Example of such worms are `Code Red' and `Slammer'.
\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the `Code Red II' and `Nimda' worm.
\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
%\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.  This is done to speed up the spread of worms at the initial stage. --This list consists of potentially vulnerable machines that are gathered beforehand and targeted first when the worm is released. An extreme case for the hit-list-scanning worms is a flash worm, which gathers all vulnerable machines into the list.--
\item \textbf{Sequential scanning:}
With sequential scanning, the worm will scan the IP addresses sequentially. This means that once a vulnerable host is compromised, it will look for IP addresses that are near to this host. For example, the address of the host is A, the next addresses that the worm will scan are A+1, or A-1. This method is used by the `Blaster worm'.
\item \textbf{DNS random scanning:} Another strategy is a kind of strategy in which the DNS infrastructure is used to locate a new target address. The IP address table from a DNS server is acquired from DNS records. The speed of a DNS scanning worm in the IPv6 internet is comparable to the speed of an IPv4 random scanning worm. There are some difficulties with this scanning technique. First it is difficult to obtain the whole address space from the DNS records. Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big. Last, the IP addresses stored in the address table are only hosts with public domain names.
\item \textbf{Routable scanning:} The worms using routable scanning acquire target IP addresses based on the routing information in a network. Through  the BGP routing tables they can scan the routable address space. This method is three times faster than a traditional worm that uses random scanning. Examples are `Spyb0t' or `network Bluepill'. 
%--a ''routing worm'' [20]. Zou et al. designed two types of routing worms [20]. One type, based on Class-A (x.0.0.0/8) address allocations, is thus called 'ClassA routing worms.'' Such worms can reduce the scanning space to 45.3\% of the entire IPv4 address space. The other type, based on BGP routing tables, is thus called''BGP routing worms'' Such worms can reduce the scanning space to only about 28.6\% of the entire IPv4 address space.--
%\end{itemize}

\item \textbf{Email and other client application worms:} An email worms uses the email systems to find email addresses to propagate. Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
file sharing systems, which have been used by worms to propagate in a similar way as email worms. For example, the Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express. % Another example is `GhostNet'

@@

Kind of network model: 
Kind of model: SIS, SIR

\subsubsection{Simple Epidemic model}

\subsubsection{Kermack-Mckendrick model}

\subsubsection*{SIS}

\subsubsection*{SIR}

\subsubsection{Two-Factor model}

\subsubsection*{self disciplinary worms}
Model for self disciplinary worms and counter measures ... []

Popular mechanism that worms use to detect vulnerable targets by random ip scanning probing. Feasible due to use 32-bit addresses. 128-bit addresses life harder for worms, except the ones that use email systems to propagate. two new strategies: uniformly distributed random number generator to select new target. :spread locally, by biasing the search space towards addresses within the same subnet or network.  
The second strategy is almost the same as what email virusses would get. For this reason we work an example out ..

>>>>

Infection by worms start with dropping the work on the network. Different dropping mechanisms can be used, whereby the initial attack can be random or targeted at a specific node. Frequently used mechanisms are USB sticks (given to a specific person or left behind for pick up by a random person), email, or malicious software through fishing or trojan horses. More important for determining our delay is the propagation strategy. Propagation is performed by determining the next nodes to spread the worm to. The following are common propagation methods: 
\begin{description}
%\item Scanning methods: A worm tries to guess (scan) the address of potential target nodes to infect. It can use distinct ways of scanning.
%% random, localized, topological or hit-list scanning.
%\begin{itemize}
\item \textbf{Selective Random scanning:} Some worms propagate by using the method of selective random targeting IP addresses. The worm selects randomly a part of the selected address space instead of scanning the whole address space. The reserved address blocks and the unassigned addresses are excluded from the address space. The rate of success for randomly chosen IP addresses is very low, but it is easy to implement. Example of such worms are `Code Red' and `Slammer'.
\item \textbf{Localized scanning:} A worm that uses localized scanning will scan for hosts in the local address space. This method is used by the `Code Red II' and `Nimda' worm.
\item \textbf{Topological scanning:} Address information stored in the victim machines are used to locate new targets. This method used by the `Morris' worm.
%\item \textbf{Hit-list scanning:} With the method of hit-list scanning, a short-list of vulnerable systems is made beforehand.  This is done to speed up the spread of worms at the initial stage. --This list consists of potentially vulnerable machines that are gathered beforehand and targeted first when the worm is released. An extreme case for the hit-list-scanning worms is a flash worm, which gathers all vulnerable machines into the list.--
\item \textbf{Sequential scanning:}
With sequential scanning, the worm will scan the IP addresses sequentially. This means that once a vulnerable host is compromised, it will look for IP addresses that are near to this host. For example, the address of the host is A, the next addresses that the worm will scan are A+1, or A-1. This method is used by the `Blaster worm'.
\item \textbf{DNS random scanning:} Another strategy is a kind of strategy in which the DNS infrastructure is used to locate a new target address. The IP address table from a DNS server is acquired from DNS records. The speed of a DNS scanning worm in the IPv6 internet is comparable to the speed of an IPv4 random scanning worm. There are some difficulties with this scanning technique. First it is difficult to obtain the whole address space from the DNS records. Secondly, the worm has to carry the database, which can slow down the propagation if the database is to big. Last, the IP addresses stored in the address table are only hosts with public domain names.
\item \textbf{Routable scanning:} The worms using routable scanning acquire target IP addresses based on the routing information in a network. Through  the BGP routing tables they can scan the routable address space. This method is three times faster than a traditional worm that uses random scanning. Examples are `Spyb0t' or `network Bluepill'. 
%--a ''routing worm'' [20]. Zou et al. designed two types of routing worms [20]. One type, based on Class-A (x.0.0.0/8) address allocations, is thus called 'ClassA routing worms.'' Such worms can reduce the scanning space to 45.3\% of the entire IPv4 address space. The other type, based on BGP routing tables, is thus called''BGP routing worms'' Such worms can reduce the scanning space to only about 28.6\% of the entire IPv4 address space.--
%\end{itemize}

\item \textbf{Topology-based Wroms} Email and other client application worms:An email worms uses the email systems to find email addresses to propagate. Other client applications can include: Internet Relay Chat (IRC), Instant Messenger (IM), and a variety of peer-to-peer
file sharing systems, which have been used by worms to propagate in a similar way as email worms. For example, the Kak worm is a Javascript computer worm that spread itself by exploiting a bug in Outlook Express. % Another example is `GhostNet'

@@

SI, SIR, SIS assume $\beta$ to be constant.

Continious-time model
\subsubsection{Simple Epidemic model}
%Codered paper gebruikt voor model
A Simple epidemic model is SI model. Nodes in the network can be either susceptible or infected.
Once a node is infected it will stay infected. 
The simple epidemic model is considered to be of a fixed size. The model for a fixed population is as follows:
\begin{equation}
\dfrac{d I(t)}{dt} = \beta I(t)[N-I(t)]
\end{equation} 
where I(t) is the number of infected hosts at time t; $\beta$ is the propagation rate; and N is the number of nodes in the network. In the beginning, $t=0$, I(0) hosts are infected. All the other nodes, $N - I(0)$, in the network are susceptible. 
The solution of this equation is the following logistic curve:
\begin{equation}
I = \dfrac{e^{\beta(t-T)}}{1+e^{\beta(t-T)}}
\end{equation}
where T is a time parameter representing the point of maximum increase in the
growth.\\

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.5]{Images/SEMmodel.png}
\caption{Simple Epidemic Model. The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network. Original image from \cite{OnWorms2005survey}}
\label{SEMmodel}
\end{figure}

Figuur \ref{SEMmodel} laat het SEM model zien.
This Simple epidemic model has been used in various papers \cite{OwnInternetSI} \cite{CodeRed} to model random scanning worm such as Code Red and Slammer. \\
For our paper we have to approximate the propagation speed en Infect gelijkstellen aan het aantal genoeg nodes die geinfecteerd moeten zijn door de APT om de controle te hebben. Dan weten we hoelang het duurt. In het begin zullen er ook geen geinfecteerde hosts zijn, dus de vergelijking kan opgelost worden.\\
Er wordt geen rekening gehouden met de topology ? nog uitzoeken.
\subsubsection{Kermack-Mckendrick model: SIR}
In the Kermack-Mckendrick model, also known as SIR model, nodes can have three states: susceptible, infected or removed. Once a node of the network has been recovered from a worm, the node will stay in the removed mode and never becomes infected again. These nodes are not able to infect other nodes and can no longer be infected. \\
Let I(t) be the number of infectious hosts at time \textit{t}, R(t) be the number of removed hosts at time \textit{t} and J(t) is the number of infected hosts by time \textit{i}, regardless the fact that a node can be in a removed state.
\begin{equation}
J(t) = I(t) + R(t).
\end{equation} 
The Kermack-McKendrick model can be represented as follows:
\begin{equation}
\begin{Bmatrix} \dfrac{d J(t)}{dt} = \beta J(t) \big[N- J(t) \\
\dfrac{d R(t)}{dt} = \gamma I(t) \\
J(t) = I(t) + R(t) = N - S(t)
 \end{Bmatrix}
\end{equation}
Parameter $\beta$ is again the rate of infection and $\gamma$ is the rate of removal of infected hosts. S(T) is the number of susceptible hosts at time t.
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.5]{Images/KMmodel.png}
\caption{KM Model. The x-coordinate is the propagation time and the y-coordinate is the infected percentage of the whole network. Original image from \cite{OnWorms2005survey}. $N=10000, \beta = 1/10000000$.}
\label{KMmodel}
\end{figure}

%codered%
--Define $\rho = \gamma/\beta$ to be the relative removal rate [3]. One
interesting result coming out of this model is
dI(t)
dt > 0 if and only if S(t) > p. (6)
Since there is no new susceptible host to be generated,
the number of susceptible hosts S(t) is a monotonically decreasing
function of time t. If S(t0) < p, then S(t) < p and
dI(t)/dt < 0 for all future time t>t0. In other words, if the
initial number of susceptible hosts is smaller than some critical
value, S(0) < p, there will be no epidemic and outbreak
[15].
The Kermack-Mckendrick model improves the classical
simple epidemic model by considering that some infectious
hosts either recover or die after some time. However, this
model is still not suitable for modeling Internet worm propagation.
First, in the Internet, cleaning, patching, and filtering
countermeasures against worms will remove both susceptible
hosts and infectious hosts from circulation, but KermackMckendrick
model only accounts for the removal of infectious
hosts. Second, this model assumes the infection rate
to be constant, which isn't true for a rampantly spreading
Internet worm such as the Code Red worm--
\subsubsection*{SIS}

\subsubsection{Two-Factor model}
%distributed worm simulation
--Zou et
al. [15] present a ''two-factor'' model that extends SIR
epidemiological model to capture the effects of human
countermeasures and the congestion due to the worm
spread. Shen et al. [16] provide a discrete-time worm
model that considers patching and cleaning effect and can
model worms with local scanning techniques. All these
approaches abstract specifics of the Internet topology,
change in the size of vulnerable population as the worm
spreads, and the effect of the individual host and network
defenses (except for patching, in case of [16]) on the
spread--
\subsubsection*{self disciplinary worms}
Model for self disciplinary worms and counter measures ... []

Popular mechanism that worms use to detect vulnerable targets by random IP scanning probing. Feasible due to use 32-bit addresses. 128-bit addresses life harder for worms, except the ones that use email systems to propagate. two new strategies: uniformly distributed random number generator to select new target. :spread locally, by biasing the search space towards addresses within the same subnet or network.  
The second strategy is almost the same as what email virusses would get. For this reason we work an example out ..
